First Exclusive Measurement of Deeply Virtual Compton Scattering off $^4$He: Toward the 3D Tomography of Nuclei

### 2018-01-08 eg6-letters/coherent/dvcs_he4_prl.tex

%\documentclass[preprint,nofootinbib,showpacs,prl,superscriptaddress,secnumarabic,amssymb,nobibnotes,aps,floatfix]{revtex4}
\documentclass[twocolumn,nofootinbib,showpacs,prl,superscriptaddress,secnumarabic,amssymb,nobibnotes,aps,floatfix]{revtex4}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
%
\def\Dirac#1{#1\hskip-5pt/}
%

@@

\input author_list.tex
%
\date{\today}
\begin{abstract}
We report on the first measurement of the beam-spin asymmetry in the exclusive 
process of coherent deeply virtual Compton scattering off a nucleus. The 
experiment used the 6 GeV electron beam from the CEBAF accelerator at Jefferson 
Lab incident on a pressurized $^4$He gaseous target placed in front of the CEBAF Large Acceptance 
Spectrometer (CLAS). The scattered electron was detected by CLAS and the photon 
by a dedicated electromagnetic calorimeter at forward angles. To ensure the 
exclusivity of the process, a specially designed radial time projection chamber 
was used to detect the recoiling $^4$He nuclei. We measured beam-spin 
asymmetries larger than those observed on the free proton in the same kinematic 
domain. From these, we were able to extract, in a model-independent way, the 
real and imaginary parts of the only $^4$He Compton form factor, $\cal H_A$. 
This first measurement of coherent deeply virtual Compton scattering on 
the $^4$He nucleus, with a fully exclusive final state via nuclear recoil 
tagging, leads the way toward 3D imaging of the partonic structure of nuclei.
\end{abstract}
\pacs{Valid PACS appear here}

@@

The generalized parton distribution (GPD) framework offers the opportunity to 
obtain information about the momentum and spatial degrees of freedom of the 
quarks and gluons inside hadrons \cite{Mueller:1998fv,Ji:1996ek,Ji:1996nm,
Radyushkin:1996nd,Radyushkin:1997ki}. In impact parameter space the GPDs are 
indeed interpreted as a tomography of the transverse plane for partons 
carrying a given fraction of the proton longitudinal momentum 
\cite{Burkardt:2000za,Diehl:2002he,Belitsky:2002ep,Burkardt:2005hp}. The most 
promising way to access GPDs experimentally is through the measurement of 
deeply virtual Compton scattering (DVCS), \textit{i.e.}, the hard exclusive 
electroproduction of a real photon on a hadron. While other processes are also known 
to be sensitive to GPDs, the measurement of DVCS is considered the cleanest 
probe and has been the focus of efforts at Jefferson Lab, HERA, and CERN 
\cite{Stepanyan:2001sm,Airapetian,Airapetian:2010nu,Chekanov:2003ya,Adloff:2001cn,Aktas:2005ty,Chen:2006na,
MunozCamacho:2006hx,Girod:2007aa,Mazouz:2007aa,Gavalian:2009,Seder:2015,Defurne:2015kxq,
Pisano:2015,Jo:2015ema,Joerg:2016hhs}. The vast majority of these measurements 
focused on the study of the proton and allowed for an extraction of its 
three-dimensional image (for reviews of the field, see~\cite{Goeke:2001tz,
Diehl:2003ny,Ji:2004gf,Belitsky:2005qn,Boffi:2007yc,Guidal:2013rya}). These 
recent developments could also be applicable to nuclei, giving access to novel 
information about nuclear structure in terms of quarks and gluons~\cite{Berger:2001zb,
Cano:2003ju,Guzey:2005ba,Dupre:2015jha}. Such a study of the 3D 
structure of nuclei appears to be especially interesting in light of the large, 
yet unresolved, nuclear effects observed in nuclear parton distribution 
functions~\cite{Geesaman:1995yd,Norton:2003cb,Hen:2016kwk}. The results 
presented in this letter demonstrate the feasibility of such an approach and 
constitute the first step toward a tomography of nuclei.

@@

Figure~\ref{fig:diags} illustrates the handbag diagram for coherent DVCS on 
$^4$He, where the four-vectors of the electrons, photons, and $^4$He are denoted 
by $k/k^\prime$, $q/q^\prime$, and $p/p^\prime$ respectively. For large 
virtual photon 4-momentum squared, $Q^2=-(k-k')^{2}$, and small squared 
momentum transfer, $t=(p-p')^{2}$, the DVCS handbag 
diagram can be factorized into two parts \cite{Freund_Collins,Ji_Osborne}. The 
hard part includes the photon-quark interaction and is calculable in 
perturbative QED. The non-perturbative part is parametrized in terms of GPDs, 
which embed the partonic structure of the hadron. The GPDs depend on the three 
variables $x$, $\xi$, and $t$. $\xi$ relates to the Bjorken variable 
$x_{A}$: $\xi\approx {{x_A}\over{2-x_A}}$, where $x_A=\frac{Q^2}{2M_A\nu}$, 
$\nu$ is the energy of the virtual photon, and 
$M_A$ is the nuclei mass. $x$ is the quark's internal loop momentum 
fraction and cannot be accessed experimentally in DVCS. We in fact measure 
Compton form factors (CFF), which are complex quantities defined as:
\begin{align}
\begin{split}
\Re e(&\mathcal{H}_{A}) = \\
    &\mathcal{P} 
\int_{0}^{1}dx[H_A(x,\xi,t)-H_A(-x,\xi,t)] \, C^{+}(x,\xi), 
\end{split} \\
\Im m(&\mathcal{H}_{A}) = - \pi ( H_A(\xi,\xi,t)-H_A(-\xi,\xi,t)),
\end{align}
with $H_A$ a GPD, $\mathcal{P}$ 
the Cauchy principal value integral, and a coefficient function $C^{+}= 
\frac{1}{x-\xi} + \frac{1}{x+\xi}$.

Until now, the only available data on nuclear DVCS were from the HERMES 
experiment \cite{Airapetian:2010nu}. In this experiment, the exclusivity of 
the reaction was obtained through kinematic cuts using only the measured 
scattered electron and real photon. This measurement was performed on a large 
set of nuclei ($^4$He, $^{14}$N, $^{20}$Ne, $^{85}$Kr and $^{131}$Xe), 
but contamination from incoherent processes can be suspected to influence the results 
significantly~\cite{Guzey:2003jh}. The direct detection of the recoil nucleus 
can however guarantee that the nucleus remains intact.

The $^4$He nucleus is an ideal experimental target for nuclear DVCS, as it is 
light enough to be detected by our experimental setting, while it is subject 
to interesting nuclear effects~\cite{JSeely}. Its zero spin also leads to an 
important simplification, as a spin-zero hadron is parametrized by only one 
chiral even GPD ($H_{A}(x,\xi,t)$) at leading-twist, while four GPDs arise for 
the spin-$\frac{1}{2}$ nucleon. This significantly simplifies the interpretation of the 
data and allows a model independent extraction of the $^4$He CFF 
($\mathcal{H}_{A}$) presented at the end of this letter.

The CEBAF Large Acceptance Spectrometer (CLAS) in Hall B at Jefferson 
Lab~\cite{Mecking:2003zu} has been previously supplemented with an inner 
calorimeter (IC) and a solenoid magnet to measure DVCS observables on the nucleon
\cite{Girod:2007aa,Gavalian:2009,Seder:2015,Pisano:2015,Jo:2015ema}. The IC 
extended the photon detection acceptance of CLAS to polar angles as low as 
4$^{\circ}$. The 5-T solenoid magnet acted as a guiding field for the 
low-energy M\o{}ller electrons that were absorbed in a  
heavy shield placed around the beam line.

In the kinematic range of the present experiment, the recoil $^4$He nuclei 
have low momentum, averaging 300 MeV. CLAS could not detect such low-energy $\alpha$ 
particles, so in order to ensure the exclusivity of the measurement, we built a 
small and light radial time projection chamber (RTPC). The 
RTPC was a $20$-cm-long cylinder with a diameter of $15$ cm, positioned in the 
solenoid magnet. In the center of the RTPC was the target cell, a $25$-cm-long 
and $6$-mm-diameter Kapton tube with $27$-$\mu$m-thick walls filled with 
gaseous $^4$He at 6~atm (see~\cite{Dupre:2017upj} for a detailed description of 
the RTPC and its performances). The experiment (E08-024)~\cite{Hafidi:2008pr} collected
data over 40 days at the end of 2009 using a 
nearly 100\% duty factor, longitudinally polarized electron beam (83.7$\pm 3.5 \%$ 
polarization~\cite{Perrin:thesis}) at an energy of 6.064 GeV. The RTPC was calibrated specifically 
for the detection of $^4$He nuclei using elastic scattering 
($e^4$He $\to e^\prime$$^4$He$'$) with a 1.2~GeV electron beam.

%DVCS selection
To identify coherent DVCS events, we first selected events where one electron, 
one $^4$He, and at least one photon were detected in the final state. Electrons 
were identified using their measured momentum, light yield, time, and energy obtained from 
the CLAS drift chambers, \v{C}erenkov counters, scintillator counters, 
and electromagnetic calorimeters, respectively. The recoiling $^4$He nuclei 
were identified in the RTPC using time and energy-loss cuts for tracks in the 
fiducial region~\cite{Hattawy:thesis}. In addition, we applied a vertex-matching 
cut to ensure that the 
electron and helium nucleus originated from a common reaction vertex in the 
target cell. The photons were 
detected in either the IC or the CLAS electromagnetic calorimeters. Note that 
even though the DVCS reaction has only one real photon in the final state, 
events with more than one good photon were not discarded at this stage. These
were mainly caused by accidental coincidences of soft photons and did not directly
affect this measurement, as only the most energetic photon of an 
event was considered a DVCS photon candidate.
This prescription however slightly increased the corrections associated with
the $\pi^0$ and the accidental backgrounds discussed below.

We selected events with $Q^{2}$ greater than 1~GeV$^{2}$ for which the DVCS handbag 
diagram is believed to be dominant. Then the exclusivity of the reaction was ensured by 
applying a set of cuts on the following kinematic variables: the coplanarity 
angle $\Delta\phi$ between the ($\gamma,\gamma^*$) 
and ($\gamma^*$,$^4$He$'$) planes, the missing energy, mass, and transverse
momentum of the $e'^4$He$'\gamma$ system, the missing mass squared of 
the $e'^4$He$'$ system, and the angle $\theta$
between the measured photon and the missing momentum of the $e'^4$He$'$ system.  
The experimental data for the most relevant exclusivity variables and applied
cuts are shown in Fig.~\ref{fig:kin-cuts} (see \cite{Hattawy:thesis} for 
additional details). We also 
rejected events where a $\pi^0$ was identified by the invariant mass of 
two photons. About 3200 events passed all these requirements, their kinematic 
distributions\footnote{We use here and for other results 
$x_B =\frac{Q^2}{2M_N\nu}$ with $M_N$ the proton mass instead of $x_A$.
This makes it easier to compare these results with the proton DVCS data
available in the literature.} are shown in Fig.~\ref{fig:kin-coverage}.

\begin{figure}[tb]
\centering
\hspace{-0.45cm}
\includegraphics[width=9cm]{F_all_coh_exc_cuts-pub.pdf}
\caption{Four of the six coherent DVCS exclusivity cuts. The black 
distributions represent the initial candidate events, while the shaded 
distributions represent those that passed all of the exclusivity cuts except 
the quantity plotted. The vertical red lines represent the applied cuts.
The distributions from left to right and from top to bottom are: coplanarity 
angle $\Delta \phi$, missing energy $E_X$, missing-mass-squared $M_X^2$, and 
the cone angle $\theta$ between the measured photon and the missing momentum 
of the $e^\prime{^4}$He$^\prime$ system.}
\label{fig:kin-cuts}
\end{figure}

@@

We identified two main backgrounds, accidental coincidences and exclusive coherent
$\pi^0$ production. The accidentals have particles originating from different events,
and we estimated their contribution to be 4.1\% of the data sample. We evaluated this 
contribution by selecting events passing all the cuts but with the scattered electron and 
$^4$He originating from different vertices. The $\pi^0$ production can
be mistaken for DVCS when one of the two photons from the $\pi^0$ 
decay is produced at low-energy in the laboratory frame and remains undetected.  
To estimate the effect of this contamination, we developed an event generator 
tuned on the experimental yield of exclusive $\pi^0$ with two photons measured. 
We used this generator together with a GEANT3 simulation of our 
detectors to estimate the ratio of the number of $\pi^0$ events where the 
two photons are detected to those that are misidentified as DVCS events. This 
ratio is then multiplied by the measured yield of exclusive $\pi^0$ events to 
correct the DVCS data. Depending on the kinematics, we found contaminations of 
2 to 4\%.

%asymmetries
In this work, the physics observable extracted using coherent DVCS events is
the beam-spin asymmetry, $A_{LU}$. On an unpolarized target, $A_{LU}$ is 
defined as the difference of cross sections for the reaction with opposite beam 
helicities normalized to the total cross section:
  \begin{equation}
  A_{LU} = \frac{d^{4}\sigma^{+} - d^{4}\sigma^{-} }
                {d^{4}\sigma^{+} + d^{4}\sigma^{-}},
    \label{BSA_equation}
  \end{equation}
where $d^{4}\sigma^{\pm}$ is the DVCS differential cross 
section for positive (negative) beam helicity.

In this ratio, luminosity normalization and
detector efficiencies largely cancel and $A_{LU}$ can be 
extracted from the reaction yields for the two helicities ($N^{\pm}$):
\begin{equation}
A_{LU} = \frac{1}{P_{B}} \frac{N^{+} - N^{-}}{N^{+} + N^{-} },
\end{equation}
where $P_{B}$ is the degree of longitudinal polarization of the incident electron beam.

There is an additional process contributing to the same final state as the 
DVCS, the so-called Bethe-Heitler (BH) process, where the real photon is 
emitted by the incoming or the outgoing lepton. The DVCS and BH processes are 
indistinguishable experimentally and the amplitude of the electroproduction of 
a real photon includes a sum of the amplitudes of these two processes. The BH 
amplitude depends on the target elastic form factors, which are well known in 
this kinematic region, while the DVCS amplitude depends on the GPDs. In our 
kinematics, the cross section of the real photon electroproduction is dominated 
by BH contribution, which varies strongly with $\phi$, the azimuthal angle 
between the $(e,e^\prime)$ and $(\gamma^*,^4$He$^\prime)$ planes. The DVCS 
contribution is smaller by about a factor of two, but independent of $\phi$ at twist-2~%
\cite{Belitsky:2001ns}, and thus it is possible to separate these contributions 
with a cross section measurement. However, the DVCS-BH interference term offers 
an easier experimental access by generating spin asymmetries. We have, in 
particular, for a spin-zero target the beam spin asymmetry, $A_{LU}$, which 
can be expressed at leading-order and leading-twist~\cite{Kirchner:2003wt,
Belitsky:2008bz} as
\begin{equation}
\begin{split}
&A_{LU}(\phi) = \\
&\frac{\alpha_{0}(\phi) \, \Im m(\mathcal{H}_{A})}
{\alpha_{1}(\phi) + \alpha_{2}(\phi) \, \Re e(\mathcal{H}_{A}) + \alpha_{3}(\phi) \, 
\big( \Re e(\mathcal{H}_{A})^{2} + \Im m(\mathcal{H}_{A})^{2} \big)}.
\end{split}
\label{eq:A_LU-coh}
\end{equation}
Explicit expressions of the kinematic factors $\alpha_i$ are derived from 
expressions in Ref.~\cite{Belitsky:2008bz}:
\begin{equation}
\begin{split}
&\alpha_0 (\phi)  = \frac{x_{A}(1+\epsilon^2)^2}{y} S_{++}(1) \sin(\phi) \\
&\alpha_1 (\phi)  =  c_0^{BH}+c_1^{BH} \cos({\phi})+c_2^{BH} \cos(2\phi) \\
&\alpha_2 (\phi)  =  \frac{x_{A}(1+\epsilon^2)^2}{y}  \left( C_{++}(0) +  
C_{++}(1) cos(\phi) \right) \\
&\alpha_3 (\phi) = \frac{x^{2}_{A}t(1+\epsilon^2)^2}{y} {\mathcal P}_1(\phi) 
{\mathcal P}_2(\phi) \cdot 2 \frac{2-2y+y^2 + \frac{\epsilon^2}{2}y^2}{1 + 
\epsilon^2},
\end{split}
\end{equation}
where $S_{++}(1)$ and $C_{++}(0,1)$ are the Fourier harmonics of the 
interference amplitude in the leptonic tensor, $c_{0,1,2}^{BH}$ 
the Fourier harmonics of the BH amplitude, and
finally ${\mathcal P}_{1,2}(\phi)$ the BH
propagators, which include $\cos(\phi)$ dependencies\footnote{The explicit 
expression of all these terms can be found in Appendix~A of \cite{Armstrong:2017wfw}.}.
We observe that using the different $\sin(\phi)$, $\cos(\phi)$ and $\cos(2\phi)$ contributions, 
one can extract unambiguously both the imaginary and real parts of $\mathcal{H}_{A}$
with a fit of the $A_{LU}(\phi)$ distribution.

We present in Fig.~\ref{fig:alu} $A_{LU}$ as a function of azimuthal angle 
$\phi$ and the kinematical variables $Q^2$, $x_B$, and $t$. Due to limited 
statistics, these latter variables are studied separately with a two-dimensional 
data binning. The curves on 
the plots are fits using the function presented in Eq.~(\ref{eq:A_LU-coh}), 
where the real and imaginary parts of the CFF $\mathcal{H}_{A}$ are the only 
free parameters. 
%Could add a sentence to mention that it is sin phi dominated as expected 
% also that assymetries are larger than protons ?

>>>>

%\documentclass[preprint,nofootinbib,showpacs,prl,superscriptaddress,secnumarabic,amssymb,nobibnotes,aps,floatfix]{revtex4}
\documentclass[twocolumn,nofootinbib,prl,superscriptaddress,secnumarabic,amssymb,nobibnotes,aps,floatfix]{revtex4}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
%
\def\Dirac#1{#1\hskip-5pt/}
%

@@

\input author_list.tex
%
%\date{\today}

\begin{abstract}
We report on the first measurement of the beam-spin asymmetry in the exclusive 
process of coherent deeply virtual Compton scattering off a nucleus. The 
experiment uses the 6 GeV electron beam from the Continuous Electron Beam 
Accelerator Facility (CEBAF) accelerator at Jefferson Lab incident on a 
pressurized $^4$He gaseous target placed in front of the CEBAF Large Acceptance 
Spectrometer (CLAS). The scattered electron is detected by CLAS and the photon 
by a dedicated electromagnetic calorimeter at forward angles. To ensure the 
exclusivity of the process, a specially designed radial time projection chamber 
is used to detect the recoiling $^4$He nuclei.  We measure beam-spin 
asymmetries larger than those observed on the free proton in the same kinematic 
domain.  From these, we are able to extract, in a model-independent way, the 
real and imaginary parts of the only $^4$He Compton form factor, $\cal H_A$.  
This first measurement of coherent deeply virtual Compton scattering on the 
$^4$He nucleus, with a fully exclusive final state via nuclear recoil tagging, 
leads the way toward 3D imaging of the partonic structure of nuclei.
\end{abstract}
%\pacs{Valid PACS appear here}

@@

The generalized parton distribution (GPD) framework offers the opportunity to 
obtain information about the momentum and spatial degrees of freedom of the 
quarks and gluons inside hadrons \cite{Mueller:1998fv,Ji:1996ek,Ji:1996nm,
Radyushkin:1996nd,Radyushkin:1997ki}. In impact parameter space the GPDs are 
indeed interpreted as a tomography of the transverse plane for partons 
carrying a given fraction of the proton longitudinal momentum 
\cite{Burkardt:2000za,Diehl:2002he,Belitsky:2002ep,Burkardt:2005hp}. The most 
promising way to access GPDs experimentally is through the measurement of 
deeply virtual Compton scattering (DVCS), i.e., the hard exclusive 
electroproduction of a real photon on a hadron. While other processes are also 
known to be sensitive to GPDs, the measurement of DVCS is considered the 
cleanest probe and has been the focus of efforts at Jefferson Lab, HERA, and 
CERN 
\cite{Stepanyan:2001sm,Airapetian,Airapetian:2010nu,Chekanov:2003ya,Adloff:2001cn,Aktas:2005ty,Chen:2006na,
MunozCamacho:2006hx,Girod:2007aa,Mazouz:2007aa,Gavalian:2009,Seder:2015,Defurne:2015kxq,
Pisano:2015,Jo:2015ema,Joerg:2016hhs}. The vast majority of these measurements 
focused on the study of the proton and allowed for an extraction of its 
three-dimensional image (for reviews of the field, see~\cite{Goeke:2001tz,
Diehl:2003ny,Ji:2004gf,Belitsky:2005qn,Boffi:2007yc,Guidal:2013rya}). These 
recent developments could also be applicable to nuclei, giving access to novel 
information about nuclear structure in terms of quarks and gluons~\cite{Berger:2001zb,
Cano:2003ju,Guzey:2005ba,Dupre:2015jha}. Such a study of the 3D 
structure of nuclei appears to be especially interesting in light of the large, 
yet unresolved, nuclear effects observed in nuclear parton distribution 
functions~\cite{Geesaman:1995yd,Norton:2003cb,Hen:2016kwk}. The results 
presented in this Letter demonstrate the feasibility of such an approach and 
constitute the first step toward a tomography of nuclei.

@@

Figure~\ref{fig:diags} illustrates the handbag diagram for coherent DVCS on 
$^4$He, where the four-vectors of the electrons, photons, and $^4$He are 
denoted by $k/k^\prime$, $q/q^\prime$, and $p/p^\prime$, respectively. For 
large virtual photon 4-momentum squared, $Q^2=-(k-k')^{2}$, and small squared 
momentum transfer, $t=(p-p')^{2}$, the DVCS handbag diagram can be factorized 
into two parts \cite{Freund_Collins,Ji_Osborne}. The hard part includes the 
photon-quark interaction and is calculable in perturbative QED. The 
nonperturbative part is parametrized in terms of GPDs, which embed the partonic 
structure of the hadron. The GPDs depend on the three variables $x$, $\xi$, and 
$t$. $\xi$ relates to the Bjorken variable $x_{A}$: $\xi\approx 
{{x_A}\over{2-x_A}}$, where $x_A=\frac{Q^2}{2M_A\nu}$, $\nu$ is the energy of 
the virtual photon, and $M_A$ is the nuclei mass. $x$ is the quark's internal 
loop momentum fraction and cannot be accessed experimentally in DVCS.  We in 
fact measure Compton form factors (CFFs), which are complex quantities defined 
as
\begin{align}
\begin{split}
\Re(&\mathcal{H}_{A}) = \\
    &\mathcal{P} \int_{0}^{1}dx[H_A(x,\xi,t)-H_A(-x,\xi,t)] \, C^{+}(x,\xi), 
\end{split} \\
\Im(&\mathcal{H}_{A}) = - \pi [H_A(\xi,\xi,t)-H_A(-\xi,\xi,t)],
\end{align}
with $H_A$ a GPD, $\mathcal{P}$ the Cauchy principal value integral, and a 
coefficient function $C^{+}= 1/(x-\xi) + 1/(x+\xi)$.

Until now, the only available data on nuclear DVCS were from the HERMES 
experiment \cite{Airapetian:2010nu}. In this experiment, the exclusivity of 
the reaction was obtained through kinematic cuts using only the measured 
scattered electron and real photon. This measurement was performed on a large 
set of nuclei ($^4$He, $^{14}$N, $^{20}$Ne, $^{85}$Kr, and $^{131}$Xe), but 
contamination from incoherent processes can be suspected to influence the 
results significantly~\cite{Guzey:2003jh}. The direct detection of the recoil 
nucleus can, however, guarantee that the nucleus remains intact.

The $^4$He nucleus is an ideal experimental target for nuclear DVCS, as it is 
light enough to be detected by our experimental setting, while it is subject 
to interesting nuclear effects~\cite{JSeely}. Its zero spin also leads to an 
important simplification, as a spin-zero hadron is parametrized by only one 
chiral even GPD [$H_{A}(x,\xi,t)$] at leading twist, while four GPDs arise for 
the spin-$\frac{1}{2}$ nucleon. This significantly simplifies the 
interpretation of the data and allows a model-independent extraction of the 
$^4$He CFF ($\mathcal{H}_{A}$) presented at the end of this Letter.

The Continuous Electron Beam Accelerator Facility (CEBAF) Large Acceptance 
Spectrometer (CLAS) in Hall B at Jefferson Lab~\cite{Mecking:2003zu} has been 
previously supplemented with an inner calorimeter (IC) and a solenoid magnet to 
measure DVCS observables on the nucleon
\cite{Girod:2007aa,Gavalian:2009,Seder:2015,Pisano:2015,Jo:2015ema}. The IC 
extended the photon detection acceptance of CLAS to polar angles as low as 
4$^{\circ}$. The 5-T solenoid magnet acted as a guiding field for the 
low-energy M\o{}ller electrons that were absorbed in a  
heavy shield placed around the beam line.

In the kinematic range of the present experiment, the recoil $^4$He nuclei 
have low momentum, averaging 300 MeV. CLAS could not detect such low-energy 
$\alpha$ particles, so, in order to ensure the exclusivity of the measurement, 
we built a small and light radial time projection chamber (RTPC). The RTPC was 
a $20$-cm-long cylinder with a diameter of $15$ cm, positioned in the solenoid 
magnet. In the center of the RTPC was the target cell, a $25$-cm-long and 
$6$-mm-diameter Kapton tube with $27$-$\mu$m-thick walls filled with gaseous 
$^4$He at 6~atm (see~\cite{Dupre:2017upj} for a detailed description of the 
RTPC and its performances). The experiment (E08-024)~\cite{Hafidi:2008pr} 
collected data over 40 days at the end of 2009 using a nearly 100\% duty 
factor, longitudinally polarized electron beam (83.7$\pm 3.5 \%$ 
polarization~\cite{Perrin:thesis}) at an energy of 6.064 GeV. The RTPC was 
calibrated specifically for the detection of $^4$He nuclei using elastic 
scattering ($e^4$He $\to e^\prime$$^4$He$'$) with a 1.2~GeV electron beam.

%DVCS selection
To identify coherent DVCS events, we first selected events where one electron, 
one $^4$He, and at least one photon were detected in the final state. Electrons 
were identified using their measured momentum, light yield, time, and energy obtained from 
the CLAS drift chambers, \v{C}erenkov counters, scintillator counters, 
and electromagnetic calorimeters, respectively. The recoiling $^4$He nuclei 
were identified in the RTPC using time and energy-loss cuts for tracks in the 
fiducial region~\cite{Hattawy:thesis}. In addition, we applied a 
vertex-matching cut to ensure that the electron and helium nucleus originated 
from a common reaction vertex in the target cell. The photons were detected in 
either the IC or the CLAS electromagnetic calorimeters. Note that even though 
the DVCS reaction has only one real photon in the final state, events with more 
than one good photon were not discarded at this stage. These
were mainly caused by accidental coincidences of soft photons and did not directly
affect this measurement, as only the most energetic photon of an event was 
considered a DVCS photon candidate. This prescription, however, slightly 
increased the corrections associated with the $\pi^0$ and the accidental 
backgrounds discussed below.

We selected events with $Q^{2}$ greater than 1~GeV$^{2}$ for which the DVCS handbag 
diagram is believed to be dominant. Then the exclusivity of the reaction was ensured by 
applying a set of cuts on the following kinematic variables: the coplanarity 
angle $\Delta\phi$ between the ($\gamma,\gamma^*$) 
and ($\gamma^*$,$^4$He$'$) planes, the missing energy, mass, and transverse
momentum of the $e'^4$He$'\gamma$ system, the missing mass squared of the 
$e'^4$He$'$ system, and the angle $\theta$ between the measured photon and the 
missing momentum of the $e'^4$He$'$ system. The experimental data for the most 
relevant exclusivity variables and applied cuts are shown in 
Fig.~\ref{fig:kin-cuts} (see \cite{Hattawy:thesis} for additional details). We 
also rejected events where a $\pi^0$ was identified by the invariant mass of 
two photons. About 3200 events passed all these requirements; their kinematic 
distributions are shown in Fig.~\ref{fig:kin-coverage}. (We use here and for 
other results $x_B =\frac{Q^2}{2M_N\nu}$ with $M_N$ the proton mass instead of 
$x_A$. This makes it easier to compare these results with the proton DVCS data
available in the literature.)

\begin{figure}[tb]
\centering
\hspace{-0.45cm}
\includegraphics[width=9cm]{F_all_coh_exc_cuts-pub.pdf}
\caption{Four of the six coherent DVCS exclusivity cuts. The black 
distributions represent the initial candidate events, while the shaded 
distributions represent those that passed all of the exclusivity cuts except 
the quantity plotted. The vertical red lines represent the applied cuts.
The distributions from left to right and from top to bottom are coplanarity 
angle $\Delta \phi$, missing energy $E_X$, missing mass squared $M_X^2$, and 
the cone angle $\theta$ between the measured photon and the missing momentum of 
the $e^\prime{^4}$He$^\prime$ system.}
\label{fig:kin-cuts}
\end{figure}

@@

We identified two main backgrounds: accidental coincidences and exclusive 
coherent $\pi^0$ production. The accidentals have particles originating from 
different events, and we estimated their contribution to be 4.1\% of the data 
sample. We evaluated this contribution by selecting events passing all the cuts 
but with the scattered electron and $^4$He originating from different vertices.  
The $\pi^0$ production can be mistaken for DVCS when one of the two photons 
from the $\pi^0$ decay is produced at low energy in the laboratory frame and 
remains undetected.  To estimate the effect of this contamination, we developed 
an event generator tuned on the experimental yield of exclusive $\pi^0$ with 
two photons measured.  We used this generator together with a GEANT3 simulation 
of our detectors to estimate the ratio of the number of $\pi^0$ events where 
the two photons are detected to those that are misidentified as DVCS events.  
This ratio is then multiplied by the measured yield of exclusive $\pi^0$ events 
to correct the DVCS data. Depending on the kinematics, we found contaminations 
of 2-4\%.

%asymmetries
In this work, the physics observable extracted using coherent DVCS events is
the beam-spin asymmetry $A_{LU}$. On an unpolarized target, $A_{LU}$ is defined 
as the difference of cross sections for the reaction with opposite beam 
helicities normalized to the total cross section:
  \begin{equation}
  A_{LU} = \frac{d^{4}\sigma^{+} - d^{4}\sigma^{-} }
                {d^{4}\sigma^{+} + d^{4}\sigma^{-}},
    \label{BSA_equation}
  \end{equation}
where $d^{4}\sigma^{\pm}$ is the DVCS differential cross section for positive 
(negative) beam helicity.

In this ratio, luminosity normalization and detector efficiencies largely 
cancel, and $A_{LU}$ can be extracted from the reaction yields for the two 
helicities ($N^{\pm}$):
\begin{equation}
A_{LU} = \frac{1}{P_{B}} \frac{N^{+} - N^{-}}{N^{+} + N^{-} },
\end{equation}
where $P_{B}$ is the degree of longitudinal polarization of the incident electron beam.

There is an additional process contributing to the same final state as the 
DVCS, the so-called Bethe-Heitler (BH) process, where the real photon is 
emitted by the incoming or the outgoing lepton. The DVCS and BH processes are 
indistinguishable experimentally, and the amplitude of the electroproduction of 
a real photon includes a sum of the amplitudes of these two processes. The BH 
amplitude depends on the target elastic form factors, which are well known in 
this kinematic region, while the DVCS amplitude depends on the GPDs. In our 
kinematics, the cross section of the real photon electroproduction is dominated 
by the BH contribution, which varies strongly with $\phi$, the azimuthal angle 
between the $(e,e^\prime)$ and $(\gamma^*,^4$He$^\prime)$ planes. The DVCS 
contribution is smaller by about a factor of 2 but independent of $\phi$ at 
twist-2~% \cite{Belitsky:2001ns}, and thus it is possible to separate these 
contributions with a cross section measurement. However, the DVCS-BH 
interference term offers an easier experimental access by generating spin 
asymmetries. We have, in particular, for a spin-zero target the beam spin 
asymmetry $A_{LU}$, which can be expressed at leading order and leading 
twist~\cite{Kirchner:2003wt,Belitsky:2008bz} as
\begin{equation}
\begin{split}
&A_{LU}(\phi) = \\
&\frac{\alpha_{0}(\phi) \, \Im (\mathcal{H}_{A})}
   {\alpha_{1}(\phi) + \alpha_{2}(\phi) \, \Re (\mathcal{H}_{A}) + 
   \alpha_{3}(\phi) \, \big[ \Re (\mathcal{H}_{A})^{2} + \Im 
(\mathcal{H}_{A})^{2} \big]}.
\end{split}
\label{eq:A_LU-coh}
\end{equation}
Explicit expressions of the kinematic factors $\alpha_i$ are derived from 
expressions in Ref.~\cite{Belitsky:2008bz}:
\begin{equation}
\begin{split}
&\alpha_0 (\phi)  = \frac{x_{A}(1+\epsilon^2)^2}{y} S_{++}(1) \sin(\phi), \\
&\alpha_1 (\phi)  =  c_0^{BH}+c_1^{BH} \cos({\phi})+c_2^{BH} \cos(2\phi), \\
&\alpha_2 (\phi)  =  \frac{x_{A}(1+\epsilon^2)^2}{y}  \left[ C_{++}(0) +  
C_{++}(1) \cos(\phi) \right], \\
&\alpha_3 (\phi) = \frac{x^{2}_{A}t(1+\epsilon^2)^2}{y} {\mathcal P}_1(\phi) 
{\mathcal P}_2(\phi) \cdot 2 \frac{2-2y+y^2 + \frac{\epsilon^2}{2}y^2}{1 + 
\epsilon^2},
\end{split}
\end{equation}
where $S_{++}(1)$ and $C_{++}(0,1)$ are the Fourier harmonics of the 
interference amplitude in the leptonic tensor, $c_{0,1,2}^{BH}$ the Fourier 
harmonics of the BH amplitude, and finally ${\mathcal P}_{1,2}(\phi)$ the BH
propagators, which include $\cos(\phi)$ dependencies. (The explicit expression 
of all these terms can be found in Appendix~A of Ref.  
\cite{Armstrong:2017wfw}.). We observe that, using the different $\sin(\phi)$, 
$\cos(\phi)$, and $\cos(2\phi)$ contributions, one can extract unambiguously 
both the imaginary and real parts of $\mathcal{H}_{A}$ with a fit of the 
$A_{LU}(\phi)$ distribution.

We present in Fig.~\ref{fig:alu} $A_{LU}$ as a function of azimuthal angle 
$\phi$ and the kinematical variables $Q^2$, $x_B$, and $t$. Because of limited 
statistics, these latter variables are studied separately with a 
two-dimensional data binning. The curves on the plots are fits using the 
function presented in Eq.~(\ref{eq:A_LU-coh}), where the real and imaginary 
parts of the CFF $\mathcal{H}_{A}$ are the only free parameters. 
%Could add a sentence to mention that it is sin phi dominated as expected 
% also that assymetries are larger than protons ?

\begin{figure}[tb]
   \centering
\includegraphics[width=8.5cm]{F_Coherent_ALU_phi.pdf}
\caption{$A_{LU}$ as a function of azimuthal angle $\phi$. Results are presented
   for different $Q^{2}$ bins (top panel), $x_{B}$ bins (middle panel), and $t$ 
   bins (bottom panel).  The error bars represent the statistical 
uncertainties. The gray bands represent the systematic uncertainties, including 
the normalization uncertainties. The red curves are the results of fits with 
Eq.~(\ref{eq:A_LU-coh}).}
\label{fig:alu}
\end{figure}
